Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv',sep='\t')
qplot(x=age,y=friend_count,data=pf)

qplot(age,friend_count,data=pf) +
  xlim(13,90)

p1 <- ggplot(aes(cyl,hwy),data=mpg) + 
  geom_point()
p2 <- ggplot(aes(cyl,hwy),data=mpg) + 
  geom_jitter()
p3 <- ggplot(aes(cyl,hwy),data=mpg) + 
  geom_jitter(width=0.5)
grid.arrange(p1,p2,p3,ncol=3)



```

***

#### What are some things that you notice right away?
Response:
Friend count is more for age below 30. There are few peak values above age 60 and 90 which might be mostly fake.
***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x=age,y=friend_count),data=pf) + 
  geom_point(alpha=1/10) +
  xlim(13,90)

install.packages("gridExtra")
library(gridExtra)
grid.arrange(p1,p2,p3,ncol=3)


ggplot(aes(x=age,y=friend_count),data=pf) + 
  geom_point(alpha=1/10,position=position_jitter(h=0)) +
  xlim(13,90) 
#+
#  coord_trans(y = "sqrt")
summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x=age,y=friend_count),data=pf) + 
  geom_jitter(alpha=1/20) +
  xlim(13,90) 
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age,y=friend_count),data=pf) + 
  geom_jitter(alpha=1/20,position=position_jitter(h=0)) +
  xlim(13,90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x=age,y=friend_count),data=pf) + 
  geom_jitter(alpha=1/20,position=position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
summary(pf$age)
```

#### What do you notice?

***

### Alpha and Jitter
Explore the relationship between friends initiated vs age

Notes:

```{r Alpha and Jitter}
ggplot(aes(x=age,y=friendships_initiated),data=pf) +
  geom_jitter(alpha=1/10,position=position_jitter(h=0)) +
  coord_trans(y='sqrt')

ggplot(aes(x=age,y=friendships_initiated),data=pf) +
  geom_point(shape='.',position=position_jitter(h=0)) +
  coord_trans(y='sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)
age_groups <- group_by(pf,age)
fc_by_age <- summarise(age_groups,mean(friend_count))
  
```

Create your plot!

```{r Conditional Means Plot}
library(dplyr)
age_groups <- group_by(pf,age)

pf.fc_by_age <- summarise(age_groups,
          friend_count_mean=mean(friend_count),
          friend_count_median=median(friend_count),
          n=n())

pf.fc_by_age <- arrange(pf.fc_by_age,age)
head(pf.fc_by_age)


pf.fc_by_age <- 
pf %>%
  group_by(age) %>%
  summarise(friend_count_mean=mean(friend_count),
            friend_count_median=median(friend_count),
            n=n()) %>%
  arrange(age)
  
head(pf.fc_by_age,20)

ggplot(aes(age,friend_count_mean),data=pf.fc_by_age) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age,y=friend_count),data=pf) +
  coord_cartesian(xlim = c(13, 70),ylim=c(0,1000)) +
  geom_point(alpha=0.05,position=position_jitter(h=0),
             color='orange') +
  #coord_trans(y='sqrt') +
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=.1),
            linetype=2,color='blue') +
   geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=.5),
            color='blue') +
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=.9),
            linetype=2,color='blue')
```

#### What are some of your observations of the plot?
Response:

Having more than 100 friends is quite rare. 

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

cor.test(pf$friend_count,pf$age,method='pearson')
with(subset(pf,age<=70),cor.test(friend_count,age,method='pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf,age<=70),cor.test(friend_count,age))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(www_likes_received,likes_received),data=pf) +
  coord_cartesian(xlim = c(0, 5000),ylim=c(0,10000)) +
  geom_point(position=position_jitter(h=0),
             color='orange')
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(www_likes_received,likes_received),data=pf) +
  geom_point() +
  xlim(0,quantile(pf$www_likes_received,0.95)) +
  ylim(0,quantile(pf$likes_received,0.95)) +
  geom_smooth(method='lm',color='red') 
  
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf,cor.test(www_likes_received,likes_received))

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(Month,Temp),data=Mitchell) +
  geom_point()

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0.02

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell,cor.test(Month,Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

