Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
list.files()
pf<-read.csv('pseudo_facebook.tsv',sep='\t')
pf<-read.delim('pseudo_facebook.tsv')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
qplot(x=dob_day,data=pf) +
  scale_x_continuous(breaks=1:31)

##ggplot equivalent
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:



### Moira's Investigation
Notes:



### Estimating Your Audience Size
Notes:



#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:
50
#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:
10

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(x=dob_day,data=pf) + 
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month,ncol=3)
##equivalent ggplot
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count,data=pf,xlim=c(0,1000))

#other way
qplot(x=friend_count,data=pf)+
  scale_x_continuous(limits=c(0,1000))
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))+
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf,!is.na(gender)), binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))+
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count,pf$gender,summary)

```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!